### 2017-01-20 paper/modes.tex
pass on section 2

This paper focuses at lower-level parallelism, which can be enabled 
through build systems and testing frameworks.  Given the proliferation
of multi-core machines it is not surprising that testing frameworks
provide today support for parallel test execution (e.g.,
JUnit~\cite{junit-org}, TestNG~\cite{testng}, and NUnit~\cite{nunit}).
Build systems wrap parallel features offered by testing frameworks to
facilitate the use of such technology~\cite{maven-surefire-plugin}.
In the following we elaborate relevant features of these components
focusing on Java, Maven, and JUnit but the discussion can be
generalized to other languages, build systems, and testing frameworks.

@@

\Seq{} is the default configuration.  In constrast to the other modes,
it does not use parallelism.  Considering use of parallelism,
\ParClassParMeth{} is the most agressive mode, with \ParClassSeqMeth{}
and \SeqClassParMeth{} lying between \Seq{} and \ParClassParMeth{}.
\ParClassSeqMeth{} and \SeqClassParMeth{} run sequentially either
classes or methods.  Intuititvely, fixing one of these dimensions
enables developers to better organize their test suites as to avoid
dataraces~\Fix{cite}.  JUnit implements these modes through a set of
customized classes (e.g., \CodeIn{ParalllelComputer} and
\CodeIn{ExecutorService}).

@@

\Fix{--------------------- estou aqui}

\Jbc{I have to understand better how Surefire executes tests. I'm
afraid it has its own wrapper classes and reuse JUnit core
functionalities. I'm saying this because JUnit does not offer a finer
control over thread numbers while this configuration is possible with
Surefire} For example, Figure~\ref{fig:surefire} illustrates how to
configure Maven to execute classes sequentially (default behavior) and
test methods in parallel. Although this configuration is defined in
the build file (\eg, \emph{pom.xml} on Maven), \Fix{these settings are
forwarded to the underlying testing framework (\eg, JUnit). In case of
JUnit,}

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    morekeywords={parallel, threadCount, perCoreThreadCount},
    deletekeywords={true}
}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <parallel>methods</parallel>
        <threadCount>2</threadCount>
        <perCoreThreadCount>true</perCoreThreadCount>
    </configuration>
</plugin>
\end{lstlisting}
\caption{\label{fig:surefire} Example of Maven Surefire configuration:
    test methods in parallel (\emph{parallel} parameter) using two
    threads (\emph{threadCount} parameter) per core (boolean flag
    \emph{perCoreThreadCount}).}
\end{figure}

\Fix{We need a par. providing rationale in to explain (points in
  favor/against) these choices.}

Build systems, such as Maven\footnote{\url{maven.apache.org}},
typically provide the option to \emph{fork} JVM instances
proportionally to available cores in the machine.  If the option
``fork JVM'' is enabled, Maven will allocate a user-provided number of
JVM instances to each core in the machine and will allocate each test
class to exactly one JVM.  Note that there is no guarantee to
uniformly distribute load to each JVM given that the build system uses
classes as the unit to specify test tasks.\Fix{we need to know how
Maven allocates test classes to JVMs}

>>>>

This paper focuses at lower-level parallelism, which can be enabled
through build systems and testing frameworks.  Given the proliferation
of multi-core machines it is not surprising that testing frameworks
provide today support for parallel test execution (e.g.,
JUnit~\cite{junit-org}, TestNG~\cite{testng}, and NUnit~\cite{nunit}).
In the following we elaborate relevant features of testing frameworks
and builds systems.  We focused on Java, Maven, and JUnit but the
discussion can be generalized to other language and tools.

@@

\Seq{} is the default configuration for JUnit.  It does not use
parallelism, in constrast to the other execution modes.  Considering
use of parallelism, \ParClassParMeth{} is the most agressive mode.
\ParClassSeqMeth{} and \SeqClassParMeth{} run sequentially either
classes or methods.  Intuititvely, fixing one of these dimensions
enables developers to better organize their test suites as to avoid
dataraces~\Fix{cite}.  JUnit implements these modes through a set of
customized classes (e.g., \CodeIn{ParallelComputer}).

@@

Build systems delegate functionalities to the testing framework they
use~\cite{maven-surefire-plugin}.  Figure~\ref{fig:surefire}
illustrates how to configure Maven for mode
\ParClassSeqMeth{}~--~classes sequentially (default) and test methods
in parallel. Although this configuration is defined in the build file
(\eg, \emph{pom.xml} on Maven), these settings are forwarded to the
underlying testing framework (in this case, JUnit).

%% \Jbc{I have to understand better how Surefire executes tests. I'm
%% afraid it has its own wrapper classes and reuse JUnit core
%% functionalities. I'm saying this because JUnit does not offer a finer
%% control over thread numbers while this configuration is possible with
%% Surefire} For example,

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    morekeywords={parallel, threadCount, perCoreThreadCount},
    deletekeywords={true}
}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <parallel>methods</parallel>
        <threadCount>2</threadCount>
        <perCoreThreadCount>true</perCoreThreadCount>
    </configuration>
</plugin>
\end{lstlisting}
\caption{\label{fig:surefire} Example of Maven configuration: test
  methods run in parallel (\CodeIn{parallel} parameter) using two
  threads (\CodeIn{threadCount} parameter) per available core (flag
  \CodeIn{perCoreThreadCount}).}
\end{figure}

\subsubsection*{Forking JVMs}~
Build systems often rovide the option to \emph{fork} JVM instances
proportionally to available cores in the machine.  For Maven, the
options \CodeIn{forkCount} and \CodeIn{reuseForks} control this
feature~\cite{fork-and-parallel-execution-maven}.  Note that Maven
allocates each test class to exactly one JVM and there is no guarantee
to uniformly distribute load to each JVM given that the build system
uses classes as the unit to specify test tasks.  Consequently, only
combinations of fork with modes \Fix{??} are valid.

%\Fix{we
%  need to know how Maven allocates test classes to JVMs}
